---
title: "Course Project 1"
author: "Ahmed Moustafa"
date: "6/27/2021"
output: html_document
---

## Loading needed libraies

```{r Loading libraies, warning=FALSE, include=FALSE}
library(tidyverse)
library(readr)
library(Hmisc)
library(ggplot2)
library(dplyr)
library(plotly)
library(hrbrthemes)
library(dplyr)
```

## Reading the data

```{r Reading the data}
my_data <- read_csv("repdata_data_activity.zip")
```

## First question: What is mean total number of steps taken per day?

#### 1. Making a histogram of the total number of steps taken each day

```{r hist 1}
sum_day <- my_data %>% group_by(date) %>% drop_na() %>%
    summarise(sum_days= sum(steps))
histogram(sum_day$sum_days,xlab = "Steps")
```



#### 2. Calculating and report the mean and median total number of steps taken per day

```{r}
sum_day %>% drop_na() %>%
    summarise(steps_mean =mean(sum_days), steps_median = median(sum_days))
```


## What is the average daily activity pattern?
#### 1. A time series plot of the 5-minute interval and the average number of steps taken averaged across all days

```{r include=FALSE}
steps_Per_Interval <- my_data %>%
    group_by(interval) %>%
    summarise(meansteps = mean(steps, na.rm = TRUE)) 

avg_interval <- steps_Per_Interval %>%
    ggplot(aes(x=interval, y=meansteps)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("Mean Steps per Interval") +
    xlab("Interval")
    theme_ipsum()
```

```{r}
ggplotly(avg_interval)
```


The daily activity pattern shows one peak at interval 835 with mean steps of 206.1698


#### 2. The 5-minute interval that contains the maximum number of steps

```{r}
steps_Per_Interval %>% drop_na() %>% filter(meansteps == max(meansteps))

```

The **835 interval** on average across all the days in the dataset, contains the maximum number of steps.

## Imputing missing values

#### 1. Calculating the total number of missing values

```{r}
NAs_percent <- mean(is.na(my_data)) * 100
print(NAs_percent)
```
The NA% in the data set is ~ 4.37%.


#### 2. Devising a strategy for filling in all of the missing values in the dataset

##### The strategy
1. We determine the columns with the NAs, in this case its only the *steps* column.
2. We make a new copy of the data set with the NAs - the column with NAs.
3. We add a new column with that contains the value of the imputed NAs column.

#### 3. Creating a new dataset that is equal to the original dataset but with the missing data filled in

```{r}
my_data_without_na <- select(my_data, - steps)
my_data_without_na$steps_without_na <- impute(my_data$steps)
View(my_data_without_na)
```

#### 4. a. Make a histogram of the total number of steps taken each day

```{r}
imputed_data_grouped_by_date <- my_data_without_na %>% group_by(date) %>%
    summarise(sum_steps = sum(steps_without_na))
histogram(imputed_data_grouped_by_date$sum_steps, xlab = "Sum of the steps by day")
```

#### 4. b. Calculati and report the mean and median total number of steps taken per day and detrmining if these values differ from the estimates from the first data set.

```{r}
imputed_data_grouped_by_date %>% summarise(mean_1 = mean(sum_steps), 
                                           median_1 = median(sum_steps))
sum_day %>% drop_na() %>%
    summarise(steps_mean =mean(sum_days), steps_median = median(sum_days))
```

By looking at both of them we find that the imputed results are *slightly less* than the original data with missing values.


#### 4. c. What is the impact of imputing missing data on the estimates of the total daily number of steps?

The imputed results are less than the original data with missing values.

#### 5. Are there differences in activity patterns between weekdays and weekends?

 Organizing the Data
```{r}
weekly_data <- my_data_without_na
weekly_data$day <- ifelse(weekdays(weekly_data$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
weekly_data$day <- as.factor(weekly_data$day)
```

```{r}
weekly_data_weekend <- weekly_data %>% filter(day=="weekend")
weekly_data_weekday <- weekly_data %>% filter(day=="weekday")

weekly_data_weekend <- weekly_data_weekend %>% group_by(interval) %>%
    summarise(avg_step= mean(steps_without_na))
weekly_data_weekend$day <- "weekend"

weekly_data_weekday <- weekly_data_weekday %>% group_by(interval) %>%
    summarise(avg_step= mean(steps_without_na))

weekly_data_weekday$day <- "weekday"
weekly_data_binded <- rbind(weekly_data_weekday, weekly_data_weekend)
weekly_data_binded$day <- as.factor(weekly_data_binded$day)
```

Ploting the data to show if there is a difference between weekday and weekend

```{r}
ggplot(data = weekly_data_binded, mapping = aes(x= interval, y=avg_step)) + 
    geom_line() + facet_grid(day~.) + xlab("Interval") +ylab("Avrage steps per  Interval")
```

Yes there is a difference between weekday and weekend.

